---
title: "SEM Final Spring 2022"
author: "Sam Neylon"
date: "5/20/2022"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: readable
    smooth_scroll: true
---

  <style>
  
  p {line-height: 2em; ;}
  
  </style>
  

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

#### Using "here" to keep project directory as working directory
#### NOTE: Rmarkdown will try to make the file it is inside the working directory, but "here" should fix it.

library(here)
library(tidyverse)
library(data.table)
library(DBI)
library(Hmisc)
library(ggplot2)
library(lavaan)
library(semPlot)
library(knitr)

```

# Introduction

Growing income inequality has become a topic of intense, and increasing, public and academic interest, particularly since the recession and economic crisis which began in 2008. Wage inequality had been a niche, but important, ongoing interest of sociologists and economists, with the debate over the growing gap between high-wage and low-wage workers heating up in the 1990s over the concept of “Skill-Biased Technological Change” (see Acemoglu and Autor 2011). Comparative databases like LIS allowed researchers, starting in the 1980s, to examine the differing effects of state policy on different forms of income inequality (see Gornick and Jantii 2013). Renewed interest in income inequality was sparked by Piketty and Saez (2006), who used fiscal data to look at both wage and wealth inequality over centuries of time (Piketty 2017). Their work was crucial in part because their use of fiscal (tax) records allowed them access to high incomes which were not top-coded. One of the drawbacks of the household surveys which are often used for inequality research is that (to maintain respondent’s anonymity) the highest wages are top-coded out of the data. Access to confidential tax records allows researchers to truly see the extent of income inequality.    

Recently, sociologists (see Tomaskovic-Devey and Avent-Holt 2019, Tomaskovic-Devey et al. 2020) have focused their attention on the organizational level as a site where inequality is produced and reproduced. After all, if wages are set at the organizational level, then the organization should be an important site for inequality researchers. There have been two major barriers to this research. The first is the issue of data access. There are major US government surveys of business organizations, typically of business “establishments,” defined as a single physical location where one type of activity predominates. These include the Current Employment Statistics (CES) which has been around since the New Deal, the more recent Occupational Employment and Wage Statistics (OEWS) which surveys business establishments on the occupations and pay of their employees, or the Economic Census, which takes place every 5 years. There are also data sources which mostly rely on administrative data, such as the Quarterly Census of Employment and Wages (QCEW), which uses Unemployment Insurance (UI) filings in each state to measure employment and wages. The issue is, these data sources are confidential, and access to them requires a lengthy application process. The second barrier is theoretical, as the economists who often dominate this field, and secure the funding and data access necessary to truly dig into these phenomena, have a simplified theory of wages which leaves little room for the organizational level. Put simply, mainstream economics sees wages as a reflection of the value that an employee brings to the firm. Consequently, most research focuses on individual characteristics as the determinant of wages, such as education (Goldin and Katz 2009). Recent movement of these barriers has changed the research terrain. First, more researchers have been getting access to confidential establishment data (Handwerker 2020, Handwerker and Dey 2018, Forsythe 2019, Barth et al. 2014). Second, economists have become more interested in “inter-establishment inequality,” or the inequality between organizations. This interest in inter-establishment inequality has come, in part, from major economists getting access to matched employer-employee IRS data, a truly magnificent resource which stretches back decades (Song et al. 2019, Bloom et al. 2018). This IRS data actually allows researchers to look at inter-firm inequality, where firms are often made up of many establishments, rather than just at single business establishments. This organizational turn has also meant a focus on “Superstar Firms,” a theory that particular firms are able to dominate their industry, pay higher wages, and force competitors to pay lower wages in order to cut costs (Autor et al. 2020).  

For this paper, I do not have access to these confidential databases, but try to draw on the history of research into industries as a way of looking at organizations. Industry data is much easier to come by, and drawing on industry-level statistics to investigate firm dynamics is a well-established practice. For example, Lin and Tomaskovic-Devey (2013) draw on the organizations literature to argue that industry is, “a technical and normative field that influences firm behavior,” and that, “an industrial investigation is likely to provide informative estimates of firm-level processes” (2013:1296). If we can establish that particular "isomorphic" (DiMaggio and Powell 1983) managerial trends or business pressures exist in one industry and not another, we can use variation in industry to explain variation in wages or other outcomes (Cobb and Lin 2017, Kristal 2013, Lin and Neely 2020, Lin and Tomaskovic-Devey 2013). With this in mind, this paper is interested in changing wage inequality within industries over time.  

The other literature I draw on is that which argues that falling density and power of unions in the US has had deleterious effects on wage inequality (Lemieux 2008, Dosi et al. 2020, for background see Freeman and Medoff 1979, 1984). In a good summary and extension of the literature, Stansbury and Summers (2020) argue that the growth of wage inequality (amongst other ills) can best be explained not by changes in education or technology, but by “declining worker power.” They argue that, since the high tide of American unionization in the Postwar period, workers have had a steadily declining ability to secure rent-sharing from the firms where they work. Stansbury and Summers define these changes in rent-sharing in three ways. First, there has been a decline in the union wage premium, concomitant with the deunionization of the American economy. Private Sector unionization in the US has gone from 1/3 of the workforce in the 1950s, to 24% in 1973, and 6% in 2019 (2020:9). Second, they note (along with many others) a decline in the “large firm wage premium,” where workers in large firms generally made more than similar workers in smaller firms. This premium is typically associated with the “internal labor market,” an organizational phenomenon where wages in large firms are set less by competition in the general labor market, and more through an internal process of negotiation, alongside norms around promotion and job tenure which large firms put in place to attract and retain the best workers. Third, they point to the decline in “industry wage differentials,” the “industry wage premium,” or what other authors call “declining inter-industry wage dispersion” (Kim and Sakamoto 2008). The decline of these three forms of rent-sharing, or wage premia, Stansbury and Summers use to illustrate how worker power, and the organizational and institutional norms which this worker power facilitated, have degraded, and have been replaced by a far more competitive labor market, which favors employers over workers and drives wage inequality.  

During this process of deunionization, by definition, a smaller and smaller proportion of the workforce are union members, and thus a smaller proportion would be affected directly by deunionization. Western and Rosenfeld (2011) argue, however, that unions were a powerful force for raising up the bottom of the wage scale for all workers. Unions also created institutional norms of fairness, which tended to compress the wage scale in large firms, by raising the wages of the lowest, while keeping wages for the highest earners in check. Western and Rosenfeld argue that these norms of fairness benefited workers far beyond the blue-collar manufacturing strongholds of postwar unionization, and were especially important in the progress made in shrinking gender and racial wage gaps. They point out that mainstream economics research on wage inequality has tended to disregard deunionization as a source of growing wage inequality, precisely because - as I examine in this paper - these researchers are focused on the individual characteristic of being a union member. Because Western and Rosenfeld examine how deunionization has hurt both union and non-union wages (through the decline in norms of equity), their findings suggest that deunionization may explain between 1/5 and 1/3 of the growth in wage inequality, “an effect comparable to the growing stratification of wages by education” (2011:513). Changes in education (combined with changes in technology) being the dominant explanation for wage inequality (Goldin and Katz 2009). Rosenfeld (2006) has even argued that the norms of fairness created by unions even raise the wages of non-union middle managers, and that de-unionization has raised the wages of top managers at the expense of middle managers.  

# Method  

A Latent Growth Model (LGM) is well suited to my research, as I am interested in how wage inequality within industries has changed over time. Following the inequality literature, I define intra-industry inequality as the ratio of the 90th percentile wage in an industry to the 10th percentile wage in a given industry, the “90/10 Ratio.” I also explored using the log of the 90/10 Ratio, as this log can be interpreted as a percentage change in the 90/10 Ratio. However, the 90/10 Ratio is typically between 3 and 0, and the log is even smaller, making estimation difficult. I also conducted analysis on the change in the average wage of each industry (adjusted for inflation), but there isn’t room for interpretation of those results here.  

My first task is to model this change, and my second task is to explain it. Following the literature on de-unionization, I decided to include measures of union density as covariates in my model.
Because I was pressed for time, I opted for the simpler conditional LGM models which use time-invariant covariates. However, since my interest was in changing union density over time, this wasn’t ideal. A model using time-varying covariates could have been better, but I think the ideal model would be a “parallel process” model, a more structural model. In a parallel process model, I would model both intra-industry wage inequality and union density as latent curves, with union density latent intercept and latent slope affecting wage inequality latent intercept and latent slope. The parallel process model is particularly interesting for my case, because the change in union density over time is a crucial variable, and an LGM could capture the nuances of this variable over time. This will be an interesting route for future research.  

# Data  

To create industry-level variables, I started with individual-level microdata from the Current Population Survey, starting in 1990 and going till 2019. I used the IPUMS (Flood et al. 2021) version of the CPS. A classic source in the income inequality literature is the Merged Outgoing Rotation Groups of the CPS (CPS MORG). The CPS is a panel survey, where households are asked additional questions when they are in the last months of their rotations. By filtering the monthly CPS data, I was able to get a large source of income data going back to 1990, which has a much larger n than the CPS ASEC data. Additionally, the MORG provides hourly wage data, which Lemieux (2006) considers the best measure of wage inequality. Crucially, the MORG includes questions about union affiliation.  

The MORG “earner study” is limited to those who are employed, are over 16, who are not self-employed, and who are non-military. For those respondents who did not have an hourly wage, I calculated an hourly wage equivalent (following the income inequality literatures), using variables on weekly earnings and usual hours worked per week. Additionally, I limited my sample to those working more than 16 hours a week (the 5th percentile in usual weekly hours), and performed some data cleaning to eliminate very high hourly earnings, caused by things like workers making more than $1000 a week and working very few hours. In the future, my data cleaning will need to be more systematic.  

From this individual-level microdata, I estimated aggregate statistics by year and industry. This allowed me to create a panel of industries over time. For each industry-year, I computed the 90th and 10th percentile wages, and divided them. Because the 90/10 Ratio only compares wages within the same year, I did not adjust for inflation. To compute union density, I calculated the weighted total employment of that industry, and the weighted employment of unionized workers (respondents who indicated they were members of unions on the CPS), dividing union employment by total employment. I transformed this variable into a percent by multiplying by 100. Because union density can be difficult to estimate for one industry in one year, depending on which households fall into the CPS, I took the average of 1990-1993 for the Time 1 union density, and the average of 2017-2019 for the ending union density.  

My intermediary data set was in “long” format, with one column for industry, another for year, and a series of variables for industry-level statistics, such as the 90/10 Ratio for the industry in that year, or the union density. For the LGM in lavaan, I changed by data into “wide” format, where each row was an industry, and a column for each year, with the measurement of the 90/10 Ratio for that year. For my time-invariant union variables, I took the union density of that industry in 1990 (Time 0), and the change in union density over the period of my study (2019 union density – 1990 union density). For my models, I grand mean centered both of these union variables, by subtracting the mean unionization in 1990 or mean change in union density, but I present the un-centered variables in the descriptive section below.    

## Years

In my initial models, I attempted to use all 30 years of data that I had. However, this number of manifest variables (while it did give me very many degrees of freedom) caused many of my models not to converge (I believe that the csv export process cut the variables off at fewer decimal points). Instead, I used every 5th year, except for 2019, which was 4 years after 2015.  

## Data Import

Because my data cleaning file was rather long, I opted to export that file and import it here. I originally exported as a .csv, but something seemed to get lost in translation which was causing my models not to converge, so I opted to export as an R object (.rds).

```{r eval = TRUE}

# Read in 90/10 Ratio File

cps_9010 <- readRDS(file = here("SEM_src/cps_9010.rds"))

# Read in Inflation-Adjusted Average Wage File

cps_9010_full <- readRDS(file = here("SEM_src/cps_9010_full.rds"))

```

# Descriptive Analysis

Each year variable is the 90/10 Ratio for that industry in that year. The variable name starts with "inq", as R column names shouldn't start with numbers.

## Correlation Matrix

```{r eval = TRUE}

cps_9010_full %>% select(inq1990, inq1995, inq2000, inq2005, inq2010, inq2015, inq2019, uni1.c, DLTuni.c) %>% 
  cor() %>% 
  kable(digits = 2)

```

## Summary Stats

### Load Long Data

I had to load the "long" version of my data to make summary statistics easier.

```{r eval = TRUE}

cpsFULL_data <- read_csv(here("SEM_src/cps_90-2019.csv"))

```

### 90/10 Ratio Across Years

```{r eval = TRUE}

cpsFULL_data %>% summarise(
  Mean = mean(ratio90_10),
  q25 = quantile(ratio90_10, probs = .25, na.rm = TRUE),
  Median = quantile(ratio90_10, probs = .50, na.rm = TRUE),
  q75 = quantile(ratio90_10, probs = .75, na.rm = TRUE),
  Variance = var(ratio90_10, na.rm = TRUE)
) %>% 
  kable(digits = 3)

```

These statistics are calculated across years, so the mean is the mean 90/10 ratio for all years and industries over 1990-2019. The total 90/10 Ratio goes from a 25th percentile of 2.961 to a 75th percentile of 3.752. This means that the 90th percentile wage earners in the 75th percentile industry are making 3.752 times more than their colleagues making the 10th percentile wage in that industry.

# Variance by Year

```{r eval = TRUE}

year_num <- as.vector(seq(1990, 2019, by=1))

var_year <- cps_9010 %>% select(-1) %>% summarise(across(everything(), list(var))) %>% 
  t() %>% 
  as.vector()

var_year <- tibble(Year = year_num, Variance = var_year)

kable(var_year, digits = 3)

```

This is the variance of the 90/10 Ratio by year, which is the variance which I am attempting to explain with my model.

# Union Statistics

## Starting Union Density

```{r eval = TRUE}

cps_DELTAunion <- readRDS(file = here("SEM_src/cps_DELTAunion.rds"))

cps_DELTAunion %>% summarise(
  Mean = mean(union1),
  q25 = quantile(union1, probs = .25, na.rm = TRUE),
  Median = quantile(union1, probs = .50, na.rm = TRUE),
  q75 = quantile(union1, probs = .75, na.rm = TRUE),
  Variance = var(union1, na.rm = TRUE)
) %>% 
  kable(digits = 3)

```

These statistics are for the union density in 1990, which is one of my time-invariant covariates. The 75th percentile industry in 1990 had 20.92% union density, while the 25th percentile industry had 3.04% union density.

## Change in Union Density

```{r eval = TRUE}

cps_DELTAunion %>% summarise(
  Mean = mean(DELTAunion),
  q25 = quantile(DELTAunion, probs = .25, na.rm = TRUE),
  Median = quantile(DELTAunion, probs = .50, na.rm = TRUE),
  q75 = quantile(DELTAunion, probs = .75, na.rm = TRUE),
  Variance = var(DELTAunion, na.rm = TRUE)
) %>% 
  kable(digits = 3)

```

The other time-invariant covariate in my model (even though it measures change over time) is change in union density from 1990 to 2019. The 25th percentile industry had a drop in union density of 10.74%, while the 75th percentile industry had a drop of 0.62%. This means that at least 75% of industries had a fall in union density during this period, with industries on average losing 6.658% union density during this time.  

# Model 1: Latent Growth Model

For my LGM, I was purely modeling the change in the 90/10 Ratio over time.

## 1.0 Null Model

I began with a Null Model, which assumed that there was no change over time, and that the average 90/10 Ratio in 1990 could explain the 90/10 Ratio in all years. This means that my latent intercept is constrained to a variance of 0. I also constrained all residual variances to be the same across the periods.

```{r eval = TRUE}

# Specify Model

model1.0 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019
  Li ~~ 0*Li

# Fix residual variances

inq1990 ~~ r*inq1990
inq1995 ~~ r*inq1995
inq2000 ~~ r*inq2000
inq2005 ~~ r*inq2005
inq2010 ~~ r*inq2010
inq2015 ~~ r*inq2015
inq2019 ~~ r*inq2019
  
'

fit1.0 <- growth(model1.0, data = cps_9010)

```


```{r eval = TRUE}

summary(fit1.0, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 1.0 Results

The Null Model fit very badly. Goodness of fit indices should be over 0.90, but CFI was 0.000 and TLI was 0.334. Badness of fit indices should be less than 0.10, but RMSEA was 0.382 and SRMR was 0.483. I believe these fit statistics make a strong case that 90/10 Ratio has a latent growth curve, and that I should model it over time with a less constrained model.    

The mean latent intercept was 3.372, meaning that there was an average 90/10 Ratio of 3.372 across this period. This was the same as the average I took in my descriptive statistics.  

Taking a look at my factors, which had constrained residual variance of 0.354, so this null model explained 64.6% of variability in 90/10 Ratio, which is a surprising amount, considering the poor model fit.  

## 1.1 Random Intercept (Constrained Residual Variances)

Next, I allowed the latent intercept to vary over time, also called a "Random Intercept." Because my residual variances are still constrained, this model assumes no change over time, but allows different 90/10 Ratios by industry to explain variation.  

```{r eval = TRUE}

# Specify Model

model1.1 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019
#  Li ~~ 0*Li

# Fix residual variances

inq1990 ~~ r*inq1990
inq1995 ~~ r*inq1995
inq2000 ~~ r*inq2000
inq2005 ~~ r*inq2005
inq2010 ~~ r*inq2010
inq2015 ~~ r*inq2015
inq2019 ~~ r*inq2019
  
'

fit1.1 <- growth(model1.1, data = cps_9010)

```


```{r eval = TRUE}

summary(fit1.1, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 1.1 Results

The fit statistics for Model 1.1 are much, much better, with a CFI of 0.749, TLI of 0.835, RMSEA of 0.190, and SRMR of 0.155. These fit statistics are still not in the commonly-accepted range, but it is remarkable how much better the model fits the data when a random intercept is included.  

The residual variance in the model is now 0.150, meaning that the model accounts for 85% of the variation in the data. This compares to 0.354 in the null model, meaning this model explains 20.4% more than the null model. 

The R-square for this model indicates that 57.7% of the total variance in the 90/10 Ratio is explained by the differences between industries.  

## 1.2 Random Intercept, Fixed Slope (Constrained Residual Variances)

Model 1.2 had a random intercept which varies, but now also includes the estimation of a second latent variable: the latent slope. This variable is estimated (as you can see below) with each prefaced by the number of years since Time 1 (1990). The mean of the latent slope is constrained to 0, but the variance of the latent slope is estimated. The covariance between latent intercept and latent slope is constrained to be 0. Again, residual variances are constrained to be equal.  

```{r eval = TRUE}

# Specify Model

model1.2 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope

  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  Ls ~ 0*1
  Ls ~~ 0*Li

# Fix residual variances

inq1990 ~~ r*inq1990
inq1995 ~~ r*inq1995
inq2000 ~~ r*inq2000
inq2005 ~~ r*inq2005
inq2010 ~~ r*inq2010
inq2015 ~~ r*inq2015
inq2019 ~~ r*inq2019
  
'

fit1.2 <- growth(model1.2, data = cps_9010)

```


```{r eval = TRUE}

summary(fit1.2, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 1.2 Results

Inclusion of a latent slope has had mixed effects on model fit when compared with Model 1.1. The CFI has gone up to 0.770 (from 0.749), the TLI has gone slightly up to 0.844 (from 0.835), and the RMSEA has gone slightly down to 0.185 (from 0.190). The SRMR has actually gone up, to 0.274 (from 0.155), indicating worse model fit.

The residual variance has shrunk to 0.133 (from 0.150), meaning that this model explains more of the variation in the data.  

## 1.3 Random Intercept, Random Slope (Constrained Residual Variances) 

In Model 1.3, I allowed both the latent intercept and latent slope to vary, but still constrained the residuals to have equal variance.  

```{r eval = TRUE}

# Specify Model

model1.3 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope

  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
#  Ls ~ 0*1
#  Ls ~~ 0*Li

# Fix residual variances

inq1990 ~~ r*inq1990
inq1995 ~~ r*inq1995
inq2000 ~~ r*inq2000
inq2005 ~~ r*inq2005
inq2010 ~~ r*inq2010
inq2015 ~~ r*inq2015
inq2019 ~~ r*inq2019
  
'

fit1.3 <- growth(model1.3, data = cps_9010)

```


```{r eval = TRUE}

summary(fit1.3, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 1.3 Results

Adding a random slope has improved model fit. CFI has gone to 0.838 (from 0.770), TLI has gone to 0.883 (from 0.844), RMSEA has gone to 0.160 (from 0.185). The change in SRMS is notable, as it had gone up in the last model, to 0.274, but is now at 0.098, which is below 0.10, and considered a good model fit.  

The improvement in model fit, while not dramatic, is good, as my theory indicated a change in wage inequality over time. If the addition of a latent slope does not improve my model, it would be some indication that there is not a latent growth curve affecting the 90/10 Ratio.  

The covariance of Li and Ls is -0.006, which indicates that industries with higher inequality at Time 1 had a slower growth in inequality. 

The intercept of the latent slope, which indicates the amount that the 90/10 Ratio changes every year, is negative, but too small to see. I can use the lavaan package to take a closer look:

```{r eval = TRUE}

EST_fit1.3 <- parameterestimates(fit1.3)

ls_fit1.3 <- as_tibble(EST_fit1.3[33,])

kable(ls_fit1.3)

```

As we can see, the intercept of Ls is -0.00003046402, a very small, but also negative, number. This would indicate that inequality actually *dropped* in this period. However, we can see from the p-value, which is 0.984, and the CI, which overlaps 0, that this intercept is not statistically significant.  

I am also interested in the variance of the slope. The variation in slope is one of the main reasons to use an LGM, which can flexibly model different growth paths for different individuals (or in this case, industries). However, the variance of the slope is too small to see. We can draw it out:

```{r eval = TRUE}

EST_fit1.3 <- parameterestimates(fit1.3)

ls_fit1.3 <- as_tibble(EST_fit1.3[23,])

kable(ls_fit1.3)

```

The variance is very small, 0.0002366036, but it is significant with a p-value < 0.000, and a CI which doesn't overlap 0.  

## 1.4 Unconstrained Model

Next, I estimated an unconstrained model, or an Unconditional LGM, where the residual variances are allowed to vary across time periods, and the mean latent slope is estimated.      

```{r eval = TRUE}

# Specify Model

model1.4 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope
  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  
'

fit1.4 <- growth(model1.4, data = cps_9010)

```


```{r eval = TRUE}

summary(fit1.4, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 1.4 Results

The full unconstrained model had improved model fit from Model 1.3, but has still not reached a goodness-of-fit above 0.90, or badness-of-fit below 0.10. The CFI has increased to 0.879 (from 0.838), the TLI has increased slightly to 0.889 (from 0.883), and the RMSEA has slightly decreased to 0.156 (from 0.160). The SRMR has actually gone from a good fit (0.098) up to 0.117.  

Because more freely estimating the latent growth curve hasn't led to much improvement in model fit, it will probably be difficult to understand the changes in inequality over time using my model.  

The intercept of the latent slope is now much bigger, and positive, at 0.001, indicating an increase of 0.1 in the 90/10 Ratio every year. However, it is not statistically significant, with a p-value of 0.349 and CI which overlaps 0.  

Now that we have freed the residuals from constraint, we should take a look at the differing amount of variance they now explain. The variances of the manifest variables go from 0.073 (2010) to 0.193 (1990), indicating that this model explains between 80-90% of the variation in the data.  

The variance of the slope is again too small to see, but is significant, with a CI which doesn't overlap 0. This means that almost all of the variation in the slope is explained by the model, but not all.  

## 1.5 Random Intercept, Fixed Slope (Unconstrained Residual Variances)

Because the latent slope intercept was so small, and not significant, and the slope variance was also very small, I thought I would try modifying my model to have a random intercept, fixed slope, but unconstrained residual variances.  

```{r eval = TRUE}

# Specify Model

model1.5 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope

  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  Ls ~ 0*1
  Ls ~~ 0*Li
'

fit1.5 <- growth(model1.5, data = cps_9010)

```


```{r eval = TRUE}

summary(fit1.5, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 1.5 Results

Unfortunately, these changes actually hurt model fit, with CFI of 0.828 (down from 0.879), TLI of 0.856 (down from 0.889), RMSEA of 0.178 (up from 0.156), and SRMR of 0.248 (up from 0.117).    
This does, however, give me confidence that I need to have a random slope in my model.  

## Modification Indices

Because my unconstrained model was still not a good fit to the data (goodness-of-fit above 0.9, badness of fit below 0.10), I wanted to see if I could improve model fit. I didn't know if it was appropriate to use modification indices to improve an LGM model (because of the different constraints than a traditional CFA or SEM), but I thought I would try.    

```{r eval = TRUE}

kable(head(as_tibble(modificationIndices(fit1.4, sort = TRUE))))

```

Unfortunately, most of these modification indices are not changes I can make. However, the top change - having the Latent intercept explain only 2019, gave me the idea to try reversing the slope constraints (see Model 1.5). The 5th suggested modification was to make the values from 2010 and 2015 covary, which I did in Model 1.7.  

## 1.6 Reverse Unconstrained Model 

For this model, the final time period (2019) is set to 0, with all other time periods set in relative positions. Otherwise, it is identical to the unconstrained model.  

```{r eval = TRUE}

# Specify Model

model1.6 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope

  Ls =~ -28*inq1990 + -24*inq1995 + -19*inq2000 + -14*inq2005 + -9*inq2010 + -4*inq2015 + 0*inq2019

'

fit1.6 <- growth(model1.6, data = cps_9010)

```


```{r eval = TRUE}

summary(fit1.6, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 1.6 Results

The reverse unconstrained model has very slightly better fit than the unconstrained model, in most cases 0.001 better fit. CFI is 0.880 (up from 0.879), TLI is 0.890 (up from 0.889), RMSEA is 0.155 (down from 0.156), and SRMR is 0.116 (down from 0.116).  

Otherwise, other parameters have not changed in notable ways.  

## 1.7 Unconstrained Modification

In Model 1.7 I followed the Modification Indices result, and set the model to estimate the covariance of 2010 and 2015.  

```{r eval = TRUE}

# Specify Model

model1.7 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope
  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  
# modification

  inq2010	~~	inq2015
  
'

fit1.7 <- growth(model1.7, data = cps_9010)

```


```{r eval = TRUE}

summary(fit1.7, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 1.7 Results

There has been a small improvement in model fit, but notable, as measures of fit are now closer to either 0.90 or 0.10. CFI is 0.896, TLI is 0.900, RMSEA is 0.148, and SRMS is 0.093.  

# Model 2: Conditional LGM

Next, I see if union density has any effect on changes in intra-industry inequality. I use two time-invariant variables. First is 'uni1.c', which is the union density (percent of employees in an industry which are members of a union) at Time 1 (1990), mean centered. Second is 'DLTuni.c', which is the delta of union density, or change in union density from 1990 to 2019 (Density in 2019 - Density in 1990), which is also mean centered.  

## 2.1 Union 1

Model 2.1 takes the unconstrained Model 1.4 and adds a covariate to explain the variation in the latent intercept and slope. This covariate is the union density of an industry in 1990.  

```{r eval = TRUE}

# Specify Model

model2.1 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope
  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  
# regression

  Li + Ls ~ uni1.c
'

fit2.1 <- growth(model2.1, data = cps_9010_full)

```


```{r eval = FALSE}

summary(fit2.1, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 2.1 Results

As a reminder, this covariate is continuous. There is a very small (< 0.000) positive effect of starting union density on the latent slope, which is statistically significant (p = 0.023). This means that industries which were more unionized in 1990 have a slightly faster rate (steeper slope) of increasing inequality.  

The coefficient of the latent intercept is -0.013, which is significant (p < 0.000). This means that industries with higher levels of unionization in 1990 have slightly lower levels of inequality in 1990.  

Fit indices indicate a good fit, but still not within the 0.90 or 0.10 range. CFI is 0.876, TLI is 0.876, RMSEA is 0.144, SRMR is 0.107.  

The variance of the latent intercept has dropped from 0.334 in Model 1.4 to 0.305 here, indicating that this model describes about 3% more variability than the unconditional model.  

## 2.2 Delta Union

Model 2.2 takes the unconstrained Model 1.4, and adds a time-invariant covariate, 'DLTuni.c', which is the mean-centered value of the change in union density during this period.  

```{r eval = TRUE}

# Specify Model

model2.2 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope
  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  
# regression

  Li + Ls ~ DLTuni.c
'

fit2.2 <- growth(model2.2, data = cps_9010_full)

```


```{r eval = FALSE}

summary(fit2.2, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 2.2 Results

The coefficient of the latent intercept is 0.022, which is significant (p < 0.000). This indicates that industries which had a decrease in unionization over this period had a higher starting level of inequality. This doesn't match up with the descriptive statistics, which generally show that unionized industries have lower levels of inequality, but this may be because I am using a variable which captures change over time as a time-invariant variable.  

The coefficient of the latent slope is very small (< 0.000), negative, but not statistically significant (p = 0.079). A negative coefficient shows that industries with decreasing unionization (a negative delta) have a faster rate of increase in inequality, but this is not statistically significant.  

Fit statistics are almost identical to Model 2.1.  

## 2.3 Two Covariates

This model includes a linear combination of both starting union density and union delta.  

```{r eval = TRUE}

# Specify Model

model2.3 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope
  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  
# regression

  Li + Ls ~ uni1.c + DLTuni.c
'

fit2.3 <- growth(model2.3, data = cps_9010_full)

```


```{r eval = FALSE}

summary(fit2.3, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 2.3 Results

All four regression coefficients were not statistically significant. In other words, the small effects we saw in Model 2.1 and Model 2.2 disappear when controlling for both. This makes sense, as these two variables are obviously closely related, and this type of time-invariant model is not the right fit for my interest in change over time.  

## 2.4 Interaction

Model 2.4 added an interaction term 'uniINT', created by multiplying uni1.c by DLTuni.c.  

```{r eval = TRUE}

# Specify Model

model2.4 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope
  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  
# regression

  Li + Ls ~ uni1.c + DLTuni.c + uniINT
'

fit2.4 <- growth(model2.4, data = cps_9010_full)

```

```{r eval = TRUE}

summary(fit2.4, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 2.4 Results

Again, none of the regression coefficients for this model were statistically significant.  

## 2.5 Seperate Regressions

This may be an illegitimate use of Conditional LGM, but I thought that starting union density would have the clearest effect on the latent intercept, while change in union density would have the clearest effect on the latent slope, so I estimated a model with two separate regressions.  

```{r eval = TRUE}

# Specify Model

model2.5 <- '
# intercept
  Li =~ 1*inq1990 + 1*inq1995 + 1*inq2000 + 1*inq2005 + 1*inq2010 + 1*inq2015 + 1*inq2019

# slope
  Ls =~ 0*inq1990 + 5*inq1995 + 10*inq2000 + 15*inq2005 + 20*inq2010 + 25*inq2015 + 29*inq2019
  
# regression

  Li ~ uni1.c
  Ls ~ DLTuni.c
'

fit2.5 <- growth(model2.5, data = cps_9010_full)

```


```{r eval = TRUE}

summary(fit2.5, rsquare = TRUE, standardized = TRUE, ci = TRUE, fit.measures = TRUE)

```

### 2.5 Results

The regression coefficient for the union density in Time 1 had a statistically significant (p = 0.002) effect on the latent intercept. The coefficient was -0.009, meaning that industries with a higher union density in 1990 had a lower level of inequality. This effect (-0.009) is slightly smaller than in Model 2.1 (-0.013), where uni1.c was the only covariate.  

The regression coefficient for the effect of change in union density on the latent slope is not statistically significant (p = 0.485). It is small and negative (< 0.000), which is similar to Model 2.2, only in Model 2.2 the p-value was much lower (p = 0.079).  

These results indicate that, not only is this specification on shaky ground methodologically, but it is also not a better or more interesting model.  

# Conclusion

Results from this analysis was mixed. Latent Growth Models may not be an appropriate way to study changes in intra-industry wage inequality, as these changes tend to be very small. There does seem to be a general trend towards greater inequality within industries, but the LGM is best suited for situations where individuals have different growth trajectories over time. LGM is a flexible method for capturing these varying growth trajectories, and also for explaining them. In my case, the differences in inequality were often small, and the trajectories were generally in the same direction. There may have been interesting variation in trajectory, but I wasn't able to see it in these initial models. In particular, the mean and variance of the latent slope was so small as to be unhelpful for analysis.  

The time-invariant approach to union density was, I think, a bust. The main result I had was that union density in 1990 (average of 1990-1993) meant lower inequality. I could have gotten this result from a descriptive analysis. Instead, the advantage of an LGM is to model and explain change over time. My other time-invariant variable wasn't really time invariant, as it tried to capture change over time with a single value. Results from the union delta were not significant, or so small as to be unimportant.  

While I don't believe that my research strategy bore fruit, I actually think an extension, or even reversal, of this strategy could be interesting. As I already mentioned, a parallel process model, where I model change in *both* inequality and union density, could be very appropriate. However, since random intercept model for inequality fit rather well, it may have been interesting to use on LGM to model the change in union density, while capturing inequality in a time-invariant, or simpler, way. In other words, to reverse the method I used here.   

Another extension which I did not attempt was to test quadratic forms of the LGM. Especially if inequality had a tendency to rise, but then to plateau at a certain point, a quadratic form may be better for capturing these trends.  

My research falls into one of the frustrating traps of inequality research, especially within industries or smaller units. While changes in inequality are very socially important, they are often small, or happen slowly over time. My models did capture an increase in inequality in this period, but the small amount of variation I was attempting to explain may have hindered how far I could take the research.  

# Works Cited

<!DOCTYPE html PUBLIC "-//W3C//DTD XHTML 1.1//EN" "http://www.w3.org/TR/xhtml11/DTD/xhtml11.dtd">
<html xmlns="http://www.w3.org/1999/xhtml" xml:lang="en">
<head>
<meta http-equiv="Content-Type" content="text/html; charset=utf-8"/>
<title>Bibliography</title>
</head>
<body>
<div class="csl-bib-body" style="line-height: 1.35; margin-left: 2em; text-indent:-2em;">
  <div class="csl-entry" style="margin-bottom: 1em;">Acemoglu, Daron, and David Autor. 2011. “Skills, Tasks and Technologies: Implications for Employment and Earnings.” Pp. 1043–1171 in <i>Handbook of Labor Economics</i>. Vol. 4. Elsevier.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=urn%3Aisbn%3A978-0-444-53452-1&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=bookitem&amp;rft.atitle=Skills%2C%20Tasks%20and%20Technologies%3A%20Implications%20for%20Employment%20and%20Earnings&amp;rft.publisher=Elsevier&amp;rft.aufirst=Daron&amp;rft.aulast=Acemoglu&amp;rft.au=Daron%20Acemoglu&amp;rft.au=David%20Autor&amp;rft.date=2011&amp;rft.pages=1043-1171&amp;rft.spage=1043&amp;rft.epage=1171&amp;rft.isbn=978-0-444-53452-1&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Autor, David, David Dorn, Lawrence F. Katz, Christina Patterson, and John Van Reenen. 2020. “The Fall of the Labor Share and the Rise of Superstar Firms*.” <i>The Quarterly Journal of Economics</i> 135(2):645–709. doi: <a href="https://doi.org/10.1093/qje/qjaa004">10.1093/qje/qjaa004</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1093%2Fqje%2Fqjaa004&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=The%20Fall%20of%20the%20Labor%20Share%20and%20the%20Rise%20of%20Superstar%20Firms*&amp;rft.jtitle=The%20Quarterly%20Journal%20of%20Economics&amp;rft.volume=135&amp;rft.issue=2&amp;rft.aufirst=David&amp;rft.aulast=Autor&amp;rft.au=David%20Autor&amp;rft.au=David%20Dorn&amp;rft.au=Lawrence%20F%20Katz&amp;rft.au=Christina%20Patterson&amp;rft.au=John%20Van%20Reenen&amp;rft.date=2020-05-01&amp;rft.pages=645-709&amp;rft.spage=645&amp;rft.epage=709&amp;rft.issn=0033-5533%2C%201531-4650&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Barth, Erling, Alex Bryson, James C. Davis, and Richard Freeman. 2014. <i>It’s Where You Work: Increases in Earnings Dispersion across Establishments and Individuals in the US</i>. National Bureau of Economic Research.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=report&amp;rft.btitle=It's%20where%20you%20work%3A%20Increases%20in%20earnings%20dispersion%20across%20establishments%20and%20individuals%20in%20the%20US&amp;rft.aufirst=Erling&amp;rft.aulast=Barth&amp;rft.au=Erling%20Barth&amp;rft.au=Alex%20Bryson&amp;rft.au=James%20C%20Davis&amp;rft.au=Richard%20Freeman&amp;rft.date=2014"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Bloom, Nicholas, Fatih Guvenen, Benjamin S. Smith, Jae Song, and Till von Wachter. 2018. “The Disappearing Large-Firm Wage Premium.” <i>AEA Papers and Proceedings</i> 108:317–22. doi: <a href="https://doi.org/10.1257/pandp.20181066">10.1257/pandp.20181066</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1257%2Fpandp.20181066&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=The%20Disappearing%20Large-Firm%20Wage%20Premium&amp;rft.jtitle=AEA%20Papers%20and%20Proceedings&amp;rft.stitle=AEA%20Papers%20and%20Proceedings&amp;rft.volume=108&amp;rft.aufirst=Nicholas&amp;rft.aulast=Bloom&amp;rft.au=Nicholas%20Bloom&amp;rft.au=Fatih%20Guvenen&amp;rft.au=Benjamin%20S.%20Smith&amp;rft.au=Jae%20Song&amp;rft.au=Till%20von%20Wachter&amp;rft.date=2018-05-01&amp;rft.pages=317-322&amp;rft.spage=317&amp;rft.epage=322&amp;rft.issn=2574-0768%2C%202574-0776&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Cobb, J. Adam, and Ken-Hou Lin. 2017. “Growing Apart: The Changing Firm-Size Wage Premium and Its Inequality Consequences.” <i>Organization Science</i> 28(3):429–46. doi: <a href="https://doi.org/10.1287/orsc.2017.1125">10.1287/orsc.2017.1125</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1287%2Forsc.2017.1125&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Growing%20Apart%3A%20The%20Changing%20Firm-Size%20Wage%20Premium%20and%20Its%20Inequality%20Consequences&amp;rft.jtitle=Organization%20Science&amp;rft.stitle=Organization%20Science&amp;rft.volume=28&amp;rft.issue=3&amp;rft.aufirst=J.%20Adam&amp;rft.aulast=Cobb&amp;rft.au=J.%20Adam%20Cobb&amp;rft.au=Ken-Hou%20Lin&amp;rft.date=2017-06&amp;rft.pages=429-446&amp;rft.spage=429&amp;rft.epage=446&amp;rft.issn=1047-7039%2C%201526-5455&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Denice, Patrick, and Jake Rosenfeld. 2018. “Unions and Nonunion Pay in the United States, 1977-2015.” <i>Sociological Science</i> 5:541–61. doi: <a href="https://doi.org/10.15195/v5.a23">10.15195/v5.a23</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.15195%2Fv5.a23&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Unions%20and%20Nonunion%20Pay%20in%20the%20United%20States%2C%201977-2015&amp;rft.jtitle=Sociological%20Science&amp;rft.stitle=SocScience&amp;rft.volume=5&amp;rft.aufirst=Patrick&amp;rft.aulast=Denice&amp;rft.au=Patrick%20Denice&amp;rft.au=Jake%20Rosenfeld&amp;rft.date=2018&amp;rft.pages=541-561&amp;rft.spage=541&amp;rft.epage=561&amp;rft.issn=23306696"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">DiMaggio, Paul J., and Walter W. Powell. 1983. “The Iron Cage Revisited: Institutional Isomorphism and Collective Rationality in Organizational Fields.” <i>American Sociological Review</i> 147–60.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=The%20iron%20cage%20revisited%3A%20Institutional%20isomorphism%20and%20collective%20rationality%20in%20organizational%20fields&amp;rft.jtitle=American%20sociological%20review&amp;rft.aufirst=Paul%20J.&amp;rft.aulast=DiMaggio&amp;rft.au=Paul%20J.%20DiMaggio&amp;rft.au=Walter%20W.%20Powell&amp;rft.date=1983&amp;rft.pages=147-160&amp;rft.spage=147&amp;rft.epage=160"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Dosi, Giovanni, Richard Freeman, Marcelo Pereira, Andrea Roventini, and Maria Enrica Virgillito. 2020. <i>The Impact of Deunionization on the Growth and Dispersion of Productivity and Pay</i>. w26634. Cambridge, MA: National Bureau of Economic Research. doi: <a href="https://doi.org/10.3386/w26634">10.3386/w26634</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=report&amp;rft.btitle=The%20Impact%20of%20Deunionization%20on%20the%20Growth%20and%20Dispersion%20of%20Productivity%20and%20Pay&amp;rft.place=Cambridge%2C%20MA&amp;rft.aufirst=Giovanni&amp;rft.aulast=Dosi&amp;rft.au=Giovanni%20Dosi&amp;rft.au=Richard%20Freeman&amp;rft.au=Marcelo%20Pereira&amp;rft.au=Andrea%20Roventini&amp;rft.au=Maria%20Enrica%20Virgillito&amp;rft.date=2020-01&amp;rft.pages=w26634&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Flood, Sarah, Miriam King, Renae Rodgers, Steven Ruggles, and J. Robert Warren. 2021. “Integrated Public Use Microdata Series, Current Population Survey: Version 9.0.”</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Adc&amp;rft.type=document&amp;rft.title=Integrated%20Public%20Use%20Microdata%20Series%2C%20Current%20Population%20Survey%3A%20Version%209.0&amp;rft.publisher=Minneapolis%2C%20MN%3A%20IPUMS&amp;rft.description=IPUMS-CPS%20is%20dedicated%20to%20integrating%20and%20disseminating%20microdata%20from%20the%20Current%20Population%20Survey%20(CPS).%20The%20CPS%20is%20a%20monthly%20U.S.%20household%20survey%20conducted%20jointly%20by%20the%20U.S.%20Census%20Bureau%20and%20the%20Bureau%20of%20Labor%20Statistics.%20IPUMS-CPS%20contains%20harmonized%20data%20from%20every%20Annual%20Social%20and%20Economic%20Supplement%20(ASEC)%20supplement%20from%201962%20to%20the%20present%20and%20all%20other%20supplements%20from%201976%20to%20present.%20In%20addition%20to%20the%20many%20of%20the%20variables%20included%20in%20census%20data%20(race%20and%20ethnicity%2C%20education%2C%20income%2C%20occupation%2C%20industry%2C%20migration%2C%20disability%2C%20and%20veteran%20status)%2C%20IPUMS-CPS%20includes%20additional%20information%20on%20work%20and%20income%2C%20such%20as%20participation%20in%20welfare%20programs%2C%20job%20searching%2C%20barriers%20to%20looking%20for%20work%2C%20union%20membership%2C%20hourly%20wages%2C%20health%20insurance%2C%20and%20tax%20liability.%20In%20addition%2C%20monthly%20supplements%20supplemental%20for%20certain%20years%20include%20information%20on%20computer%20and%20Internet%20use%2C%20displaced%20workers%2C%20fertility%2C%20food%20security%2C%20job%20tenure%2C%20tobacco%20use%2C%20volunteer%20activities%2C%20voter%20registration%2C%20and%20work%20schedules.%20IPUMS-CPS%20data%20are%20available%20through%20an%20online%20data%20access%20system%20free%20of%20charge%20to%20registered%20users%20at%20cps.ipums.org.&amp;rft.identifier=https%3A%2F%2Fcps.ipums.org&amp;rft.aufirst=Sarah&amp;rft.aulast=Flood&amp;rft.au=Sarah%20Flood&amp;rft.au=Miriam%20King&amp;rft.au=Renae%20Rodgers&amp;rft.au=Steven%20Ruggles&amp;rft.au=J.%20Robert%20Warren&amp;rft.au=undefined&amp;rft.au=undefined&amp;rft.date=2021&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Forsythe, Eliza C. 2019. “The Occupational Structures of Low- and High-Wage Service Sector Establishments.” <i>Economic Development Quarterly</i> 33(2):76–91. doi: <a href="https://doi.org/10.1177/0891242419838328">10.1177/0891242419838328</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1177%2F0891242419838328&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=The%20Occupational%20Structures%20of%20Low-%20and%20High-Wage%20Service%20Sector%20Establishments&amp;rft.jtitle=Economic%20Development%20Quarterly&amp;rft.stitle=Economic%20Development%20Quarterly&amp;rft.volume=33&amp;rft.issue=2&amp;rft.aufirst=Eliza%20C.&amp;rft.aulast=Forsythe&amp;rft.au=Eliza%20C.%20Forsythe&amp;rft.date=2019-05&amp;rft.pages=76-91&amp;rft.spage=76&amp;rft.epage=91&amp;rft.issn=0891-2424%2C%201552-3543&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Freeman, Richard L., and James L. Medoff. 1979. “The Two Faces of Unionism.” <i>The Public Interest</i> 57:69.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=The%20Two%20Faces%20of%20Unionism&amp;rft.jtitle=The%20Public%20Interest&amp;rft.volume=57&amp;rft.aufirst=Richard%20L.&amp;rft.aulast=Freeman&amp;rft.au=Richard%20L.%20Freeman&amp;rft.au=James%20L.%20Medoff&amp;rft.date=1979&amp;rft.pages=69"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Freeman, Richard, and James Medoff. 1984. <i>What Do Unions Do</i>. N.Y.: Basic Books.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=book&amp;rft.btitle=What%20Do%20Unions%20Do&amp;rft.place=N.Y.&amp;rft.publisher=Basic%20Books&amp;rft.aufirst=Richard&amp;rft.aulast=Freeman&amp;rft.au=Richard%20Freeman&amp;rft.au=James%20Medoff&amp;rft.date=1984"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Goldin, Claudia Dale, and Lawrence F. Katz. 2009. <i>The Race between Education and Technology</i>. 1st Harvard University Press paperback edition. Cambridge, Massachusetts London, England: The Belknap Press of Harvard University Press.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=urn%3Aisbn%3A978-0-674-03530-0&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=book&amp;rft.btitle=The%20race%20between%20education%20and%20technology&amp;rft.place=Cambridge%2C%20Massachusetts%20London%2C%20England&amp;rft.publisher=The%20Belknap%20Press%20of%20Harvard%20University%20Press&amp;rft.edition=1st%20Harvard%20University%20Press%20paperback%20edition&amp;rft.aufirst=Claudia%20Dale&amp;rft.aulast=Goldin&amp;rft.au=Claudia%20Dale%20Goldin&amp;rft.au=Lawrence%20F.%20Katz&amp;rft.date=2009&amp;rft.tpages=488&amp;rft.isbn=978-0-674-03530-0&amp;rft.language=eng"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Gornick, Janet C., and Markus Jäntii, eds. 2013. <i>Income Inequality: Economic Disparities and the Middle Class in Affluent Countries</i>. Stanford, Calif: Stanford Univ. Press.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=urn%3Aisbn%3A978-0-8047-8675-1%20978-0-8047-9334-6%20978-0-8047-7824-4&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=book&amp;rft.btitle=Income%20inequality%3A%20economic%20disparities%20and%20the%20middle%20class%20in%20affluent%20countries&amp;rft.place=Stanford%2C%20Calif&amp;rft.publisher=Stanford%20Univ.%20Press&amp;rft.series=Studies%20in%20social%20inequality&amp;rft.aufirst=Janet%20C.&amp;rft.aulast=Gornick&amp;rft.au=Janet%20C.%20Gornick&amp;rft.au=Markus%20J%C3%A4ntii&amp;rft.date=2013&amp;rft.tpages=515&amp;rft.isbn=978-0-8047-8675-1%20978-0-8047-9334-6%20978-0-8047-7824-4&amp;rft.language=eng"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Handwerker, Elizabeth Weber. 2020. <i>Outsourcing, Occupationally Homogeneous Employers, and Growing Wage Inequality in the United States</i>. US Department of Labor, US Bureau of Labor Statistics, Office of Employment&nbsp;….</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=book&amp;rft.btitle=Outsourcing%2C%20occupationally%20homogeneous%20employers%2C%20and%20growing%20wage%20inequality%20in%20the%20United%20States&amp;rft.publisher=US%20Department%20of%20Labor%2C%20US%20Bureau%20of%20Labor%20Statistics%2C%20Office%20of%20Employment%C2%A0%E2%80%A6&amp;rft.aufirst=Elizabeth%20Weber&amp;rft.aulast=Handwerker&amp;rft.au=Elizabeth%20Weber%20Handwerker&amp;rft.date=2020"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Handwerker, Elizabeth Weber, and Matthew Dey. 2018. <i>Megafirms and Monopsonists: Not the Same Employers, Not the Same Workers</i>. BLS Working Paper Series, 2.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=report&amp;rft.btitle=Megafirms%20and%20Monopsonists%3A%20Not%20the%20same%20employers%2C%20not%20the%20same%20workers&amp;rft.aufirst=Elizabeth%20Weber&amp;rft.aulast=Handwerker&amp;rft.au=Elizabeth%20Weber%20Handwerker&amp;rft.au=Matthew%20Dey&amp;rft.date=2018"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Kim, ChangHwan, and Arthur Sakamoto. 2008. “Declining Inter-Industry Wage Dispersion in the US.” <i>Social Science Research</i> 37(4):1081–1101. doi: <a href="https://doi.org/10.1016/j.ssresearch.2008.06.001">10.1016/j.ssresearch.2008.06.001</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1016%2Fj.ssresearch.2008.06.001&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Declining%20inter-industry%20wage%20dispersion%20in%20the%20US&amp;rft.jtitle=Social%20Science%20Research&amp;rft.stitle=Social%20Science%20Research&amp;rft.volume=37&amp;rft.issue=4&amp;rft.aufirst=ChangHwan&amp;rft.aulast=Kim&amp;rft.au=ChangHwan%20Kim&amp;rft.au=Arthur%20Sakamoto&amp;rft.date=2008-12&amp;rft.pages=1081-1101&amp;rft.spage=1081&amp;rft.epage=1101&amp;rft.issn=0049089X&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Kristal, Tali. 2013. “The Capitalist Machine: Computerization, Workers’ Power, and the Decline in Labor’s Share within U.S. Industries.” <i>American Sociological Review</i> 78(3):361–89. doi: <a href="https://doi.org/10.1177/0003122413481351">10.1177/0003122413481351</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1177%2F0003122413481351&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=The%20Capitalist%20Machine%3A%20Computerization%2C%20Workers%E2%80%99%20Power%2C%20and%20the%20Decline%20in%20Labor%E2%80%99s%20Share%20within%20U.S.%20Industries&amp;rft.jtitle=American%20Sociological%20Review&amp;rft.stitle=Am%20Sociol%20Rev&amp;rft.volume=78&amp;rft.issue=3&amp;rft.aufirst=Tali&amp;rft.aulast=Kristal&amp;rft.au=Tali%20Kristal&amp;rft.date=2013-06&amp;rft.pages=361-389&amp;rft.spage=361&amp;rft.epage=389&amp;rft.issn=0003-1224%2C%201939-8271&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Lemieux, Thomas. 2006. “Increasing Residual Wage Inequality: Composition Effects, Noisy Data, or Rising Demand for Skill?” <i>The American Economic Review</i> 96(3):461–98.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Increasing%20Residual%20Wage%20Inequality%3A%20Composition%20Effects%2C%20Noisy%20Data%2C%20or%20Rising%20Demand%20for%20Skill%3F&amp;rft.jtitle=The%20American%20Economic%20Review&amp;rft.volume=96&amp;rft.issue=3&amp;rft.aufirst=Thomas&amp;rft.aulast=Lemieux&amp;rft.au=Thomas%20Lemieux&amp;rft.date=2006&amp;rft.pages=461-498&amp;rft.spage=461&amp;rft.epage=498&amp;rft.issn=00028282"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Lemieux, Thomas. 2008. “The Changing Nature of Wage Inequality.” <i>Journal of Population Economics</i> 21(1):21–48. doi: <a href="https://doi.org/10.1007/s00148-007-0169-0">10.1007/s00148-007-0169-0</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1007%2Fs00148-007-0169-0&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=The%20changing%20nature%20of%20wage%20inequality&amp;rft.jtitle=Journal%20of%20Population%20Economics&amp;rft.stitle=J%20Popul%20Econ&amp;rft.volume=21&amp;rft.issue=1&amp;rft.aufirst=Thomas&amp;rft.aulast=Lemieux&amp;rft.au=Thomas%20Lemieux&amp;rft.date=2008&amp;rft.pages=21-48&amp;rft.spage=21&amp;rft.epage=48&amp;rft.issn=0933-1433%2C%201432-1475&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Lin, Ken-Hou, and Megan Tobias Neely. 2020. <i>Divested: Inequality in the Age of Finance</i>. New York: Oxford University Press.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=urn%3Aisbn%3A978-0-19-063831-3&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=book&amp;rft.btitle=Divested%3A%20inequality%20in%20the%20age%20of%20finance&amp;rft.place=New%20York&amp;rft.publisher=Oxford%20University%20Press&amp;rft.aufirst=Ken-Hou&amp;rft.aulast=Lin&amp;rft.au=Ken-Hou%20Lin&amp;rft.au=Megan%20Tobias%20Neely&amp;rft.date=2020&amp;rft.tpages=232&amp;rft.isbn=978-0-19-063831-3"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Lin, Ken-Hou, and Donald Tomaskovic-Devey. 2013. “Financialization and U.S. Income Inequality, 1970–2008.” <i>American Journal of Sociology</i> 118(5):1284–1329. doi: <a href="https://doi.org/10.1086/669499">10.1086/669499</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1086%2F669499&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Financialization%20and%20U.S.%20Income%20Inequality%2C%201970%E2%80%932008&amp;rft.jtitle=American%20Journal%20of%20Sociology&amp;rft.stitle=American%20Journal%20of%20Sociology&amp;rft.volume=118&amp;rft.issue=5&amp;rft.aufirst=Ken-Hou&amp;rft.aulast=Lin&amp;rft.au=Ken-Hou%20Lin&amp;rft.au=Donald%20Tomaskovic-Devey&amp;rft.date=2013-03&amp;rft.pages=1284-1329&amp;rft.spage=1284&amp;rft.epage=1329&amp;rft.issn=0002-9602%2C%201537-5390&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Piketty, Thomas. 2017. <i>Capital in the Twenty-First Century</i>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=urn%3Aisbn%3A978-0-674-97985-7&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=book&amp;rft.btitle=Capital%20in%20the%20twenty-first%20century&amp;rft.aufirst=Thomas&amp;rft.aulast=Piketty&amp;rft.au=Thomas%20Piketty&amp;rft.date=2017&amp;rft.isbn=978-0-674-97985-7&amp;rft.language=English"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Piketty, Thomas, and Emmanuel Saez. 2006. “The Evolution of Top Incomes: A Historical and International Perspective.” <i>American Economic Review</i> 96(2):200–205. doi: <a href="https://doi.org/10.1257/000282806777212116">10.1257/000282806777212116</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1257%2F000282806777212116&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=The%20Evolution%20of%20Top%20Incomes%3A%20A%20Historical%20and%20International%20Perspective&amp;rft.jtitle=American%20Economic%20Review&amp;rft.stitle=American%20Economic%20Review&amp;rft.volume=96&amp;rft.issue=2&amp;rft.aufirst=Thomas&amp;rft.aulast=Piketty&amp;rft.au=Thomas%20Piketty&amp;rft.au=Emmanuel%20Saez&amp;rft.date=2006-04-01&amp;rft.pages=200-205&amp;rft.spage=200&amp;rft.epage=205&amp;rft.issn=0002-8282&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Rosenfeld, Jake. 2006. “Widening the Gap: The Effect of Declining Unionization on Managerial and Worker Pay, 1983–2000.” <i>Research in Social Stratification and Mobility</i> 24(3):223–38. doi: <a href="https://doi.org/10.1016/j.rssm.2006.01.003">10.1016/j.rssm.2006.01.003</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1016%2Fj.rssm.2006.01.003&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Widening%20the%20gap%3A%20The%20effect%20of%20declining%20unionization%20on%20managerial%20and%20worker%20pay%2C%201983%E2%80%932000&amp;rft.jtitle=Research%20in%20Social%20Stratification%20and%20Mobility&amp;rft.stitle=Research%20in%20Social%20Stratification%20and%20Mobility&amp;rft.volume=24&amp;rft.issue=3&amp;rft.aufirst=Jake&amp;rft.aulast=Rosenfeld&amp;rft.au=Jake%20Rosenfeld&amp;rft.date=2006-07&amp;rft.pages=223-238&amp;rft.spage=223&amp;rft.epage=238&amp;rft.issn=02765624&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Song, Jae, David J. Price, Fatih Guvenen, Nicholas Bloom, and Till von Wachter. 2019. “Firming Up Inequality*.” <i>The Quarterly Journal of Economics</i> 134(1):1–50. doi: <a href="https://doi.org/10.1093/qje/qjy025">10.1093/qje/qjy025</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1093%2Fqje%2Fqjy025&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Firming%20Up%20Inequality*&amp;rft.jtitle=The%20Quarterly%20Journal%20of%20Economics&amp;rft.volume=134&amp;rft.issue=1&amp;rft.aufirst=Jae&amp;rft.aulast=Song&amp;rft.au=Jae%20Song&amp;rft.au=David%20J%20Price&amp;rft.au=Fatih%20Guvenen&amp;rft.au=Nicholas%20Bloom&amp;rft.au=Till%20von%20Wachter&amp;rft.date=2019-02-01&amp;rft.pages=1-50&amp;rft.spage=1&amp;rft.epage=50&amp;rft.issn=0033-5533%2C%201531-4650&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Stansbury, Anna, and Lawrence Summers. 2020. <i>The Declining Worker Power Hypothesis: An Explanation for the Recent Evolution of the American Economy</i>. w27193. Cambridge, MA: National Bureau of Economic Research. doi: <a href="https://doi.org/10.3386/w27193">10.3386/w27193</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=report&amp;rft.btitle=The%20Declining%20Worker%20Power%20Hypothesis%3A%20An%20explanation%20for%20the%20recent%20evolution%20of%20the%20American%20economy&amp;rft.place=Cambridge%2C%20MA&amp;rft.aufirst=Anna&amp;rft.aulast=Stansbury&amp;rft.au=Anna%20Stansbury&amp;rft.au=Lawrence%20Summers&amp;rft.date=2020-05&amp;rft.pages=w27193&amp;rft.language=en"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Tomaskovic-Devey, Donald, and Dustin Robert Avent-Holt. 2019. <i>Relational Inequalities: An Organizational Approach</i>. New York: Oxford University Press.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=urn%3Aisbn%3A978-0-19-062442-2%20978-0-19-062443-9&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Abook&amp;rft.genre=book&amp;rft.btitle=Relational%20inequalities%3A%20an%20organizational%20approach&amp;rft.place=New%20York&amp;rft.publisher=Oxford%20University%20Press&amp;rft.aufirst=Donald&amp;rft.aulast=Tomaskovic-Devey&amp;rft.au=Donald%20Tomaskovic-Devey&amp;rft.au=Dustin%20Robert%20Avent-Holt&amp;rft.date=2019&amp;rft.tpages=293&amp;rft.isbn=978-0-19-062442-2%20978-0-19-062443-9"></span>
  <div class="csl-entry" style="margin-bottom: 1em;">Tomaskovic-Devey, Donald, Anthony Rainey, Dustin Avent-Holt, Nina Bandelj, István Boza, David Cort, Olivier Godechot, Gergely Hajdu, Martin Hällsten, Lasse Folke Henriksen, Are Skeie Hermansen, Feng Hou, Jiwook Jung, Aleksandra Kanjuo-Mrčela, Joe King, Naomi Kodama, Tali Kristal, Alena Křížková, Zoltán Lippényi, Silvia Maja Melzer, Eunmi Mun, Andrew Penner, Trond Petersen, Andreja Poje, Mirna Safi, Max Thaning, and Zaibu Tufail. 2020. “Rising Between-Workplace Inequalities in High-Income Countries.” <i>Proceedings of the National Academy of Sciences</i> 117(17):9277–83. doi: <a href="https://doi.org/10.1073/pnas.1918249117">10.1073/pnas.1918249117</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1073%2Fpnas.1918249117&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Rising%20between-workplace%20inequalities%20in%20high-income%20countries&amp;rft.jtitle=Proceedings%20of%20the%20National%20Academy%20of%20Sciences&amp;rft.stitle=Proc%20Natl%20Acad%20Sci%20USA&amp;rft.volume=117&amp;rft.issue=17&amp;rft.aufirst=Donald&amp;rft.aulast=Tomaskovic-Devey&amp;rft.au=Donald%20Tomaskovic-Devey&amp;rft.au=Anthony%20Rainey&amp;rft.au=Dustin%20Avent-Holt&amp;rft.au=Nina%20Bandelj&amp;rft.au=Istv%C3%A1n%20Boza&amp;rft.au=David%20Cort&amp;rft.au=Olivier%20Godechot&amp;rft.au=Gergely%20Hajdu&amp;rft.au=Martin%20H%C3%A4llsten&amp;rft.au=Lasse%20Folke%20Henriksen&amp;rft.au=Are%20Skeie%20Hermansen&amp;rft.au=Feng%20Hou&amp;rft.au=Jiwook%20Jung&amp;rft.au=Aleksandra%20Kanjuo-Mr%C4%8Dela&amp;rft.au=Joe%20King&amp;rft.au=Naomi%20Kodama&amp;rft.au=Tali%20Kristal&amp;rft.au=Alena%20K%C5%99%C3%AD%C5%BEkov%C3%A1&amp;rft.au=Zolt%C3%A1n%20Lipp%C3%A9nyi&amp;rft.au=Silvia%20Maja%20Melzer&amp;rft.au=Eunmi%20Mun&amp;rft.au=Andrew%20Penner&amp;rft.au=Trond%20Petersen&amp;rft.au=Andreja%20Poje&amp;rft.au=Mirna%20Safi&amp;rft.au=Max%20Thaning&amp;rft.au=Zaibu%20Tufail&amp;rft.date=2020-04-28&amp;rft.pages=9277-9283&amp;rft.spage=9277&amp;rft.epage=9283&amp;rft.issn=0027-8424%2C%201091-6490&amp;rft.language=en"></span>
  <div class="csl-entry">Western, Bruce, and Jake Rosenfeld. 2011. “Unions, Norms, and the Rise in U.S. Wage Inequality.” <i>American Sociological Review</i> 76(4):513–37. doi: <a href="https://doi.org/10.1177/0003122411414817">10.1177/0003122411414817</a>.</div>
  <span class="Z3988" title="url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_id=info%3Adoi%2F10.1177%2F0003122411414817&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&amp;rft.genre=article&amp;rft.atitle=Unions%2C%20Norms%2C%20and%20the%20Rise%20in%20U.S.%20Wage%20Inequality&amp;rft.jtitle=American%20Sociological%20Review&amp;rft.stitle=Am%20Sociol%20Rev&amp;rft.volume=76&amp;rft.issue=4&amp;rft.aufirst=Bruce&amp;rft.aulast=Western&amp;rft.au=Bruce%20Western&amp;rft.au=Jake%20Rosenfeld&amp;rft.date=2011-08&amp;rft.pages=513-537&amp;rft.spage=513&amp;rft.epage=537&amp;rft.issn=0003-1224%2C%201939-8271&amp;rft.language=en"></span>
</div></body>
</html>
