---
title: "NLRB Election Data"
author: "Sam Neylon"
date: '2023-01-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(lubridate)

```

# NOTES

##(01-17-23)

Import the unionwebsite cleaned data, put into ggplots.

Data: 
  * https://github.com/reuning/unionwebsite/tree/main/gen/data
  * "cleaned_data.csv"
  
##(1-18-23)

Figuring out which cities are in NYC:
  * https://www.mapcustomizer.com/map/NLRB_2019-22_NY
  * I am going to try using some of the geocode tools used by unionwebsite guy, but FOR NOW, I am sticking with NY state, just to get the charts right.

# Import

```{r}

# raw_data <- read_csv(here("data/NLRB/unionwebsite_1-17_cleaned_data.csv"),
#                     col_types = cols(.default = col_character()))

raw_data <- read_csv(here("data/NLRB/unionwebsite_1-17_cleaned_data.csv"),
                     col_types = cols(
  Case = col_character(),
  Region = col_character(),
  Case_Name = col_character(),
  Date_Filed = col_date(format = ""),
  Date_Closed = col_date(format = ""),
  City = col_character(),
  State = col_character(),
  `Unit ID` = col_character(),
  Tally_Date = col_date(format = ""),
  Num_Eligible_Voters = col_double(),
  `Void Ballots` = col_double(),
  Labor_Union = col_character(),
  Votes_For_Union = col_double(),
  `Labor Union2` = col_character(),
  `Votes for Labor Union2` = col_double(),
  `Labor Union3` = col_character(),
  `Votes for Labor Union3` = col_double(),
  Votes_Against = col_double(),
  Total_Ballots_Counted = col_double(),
  `Runoff Required` = col_character(),
  `Challenged Ballots` = col_double(),
  `Challenges are Determinative` = col_character(),
  `Union to Certify` = col_character(),
  Tally_Type = col_character(),
  Ballot_Type = col_character(),
  Reason_Closed = col_character(),
  Status = col_character(),
  `Voting Unit (Unit A)` = col_character(),
  `Voting Unit (Unit B)` = col_character(),
  `Voting Unit (Unit C)` = col_character(),
  `Voting Unit (Unit D)` = col_character(),
  Length = col_double(),
  Tally_Quarter = col_date(format = ""),
  Filed_Quarter = col_date(format = ""),
  size = col_character(),
  Unique = col_logical(),
  Case_Type = col_character(),
  Election_Data = col_character(),
  Plot_Labor_Union = col_character(),
  National = col_character(),
  National_Count = col_double(),
  Margin = col_double(),
  Union_Cer = col_character(),
  Didnt_Vote = col_double()
))

```

# Cleaning

## Select

```{r eval=TRUE}

clean_df <- raw_data %>% 
  filter(Case_Type == "RC")

# rm(raw_data)

```

## Data Fixes

FIX MONTEFIORE!
Case: 02-RC-308481
I changed this by hand based on: https://www.bxtimes.com/montefiore-physician-residents-union-cir/
To: 1200 Num_Eligible_Voters

```{r eval=TRUE}

clean_df <- clean_df %>% 
  # Clean Montefiore Number
  mutate(
    Num_Eligible_Voters = ifelse(Case %in% "02-RC-308481", 1200, Num_Eligible_Voters)
  )

```

## New Columns

```{r eval=TRUE}

clean_df <- clean_df %>% 
  mutate(
    New_Mem = case_when(
      is.na(Margin) ~ 0,
      Margin > 0.5 ~ Num_Eligible_Voters,
      Margin <= 0.5 ~ 0),
    Won = case_when(
      is.na(Margin) ~ 0,
      Margin > 0.5 ~ 1,
      Margin <= 0.5 ~ 0),
    NY = ifelse(State == "NY", 1, 0),
    #NYC = case_when(
      #City %in% c("Astoria", "Bronx", )
    )

```

# Tables

## Num Workers

```{r eval=TRUE}

graph_table <- clean_df %>% 
  filter(between(Date_Filed, as.Date('2019-01-01'),as.Date('2022-12-31'))) %>% 
  group_by(year = floor_date(Date_Filed, 'year')) %>% 
  summarise(
    USA_work = sum(Num_Eligible_Voters),
    USA_work_win = sum(New_Mem),
    NY_work = sum(Num_Eligible_Voters*NY),
    NY_work_win = sum(New_Mem*NY)
  ) %>% 
  mutate(USA_work_NOwin = USA_work - USA_work_win,
         NY_work_NOwin = NY_work - NY_work_win,
         YEAR = year(year))

longer_table <- graph_table %>% 
  select(year, USA_work_win, NY_work_win, USA_work_NOwin, NY_work_NOwin) %>% 
  pivot_longer(c(USA_work_win, NY_work_win, USA_work_NOwin, NY_work_NOwin), names_to = "cat", values_to = "Workers") %>% 
  mutate(YEAR = year(year),
         geo = case_when(
           str_detect(cat, "USA") ~ "USA",
           str_detect(cat, "NY") ~ "NY"
         ),
         Win = case_when(
           str_detect(cat, "_win") ~ "Win",
           str_detect(cat, "_NOwin") ~ "No Win"
         ),
         geo = factor(geo, levels = c("USA", "NY")),
         Win = factor(Win)
         )


all_long_table <- graph_table %>% 
  select(year, USA_work_win, NY_work_win, USA_work, NY_work) %>% 
  pivot_longer(c(USA_work_win, NY_work_win, USA_work, NY_work), names_to = "cat", values_to = "Workers") %>% 
  mutate(YEAR = year(year),
         geo = case_when(
           str_detect(cat, "USA") ~ "USA",
           str_detect(cat, "NY") ~ "NY State"
         ),
         Win = ifelse(str_detect(cat, "_win"), "Win", "All"),
         geo = factor(geo, levels = c("USA", "NY State")),
         Win = factor(Win))


```

## Election Count

```{r eval=FALSE}

graph_table_count <- clean_df %>% 
  filter(between(Date_Filed, as.Date('2019-01-01'),as.Date('2022-12-31'))) %>% 
  group_by(year = floor_date(Date_Filed, 'year')) %>% 
  summarise(
    USA_count = count(Num_Eligible_Voters),
    USA_count_win = count(New_Mem),
    NY_count = count(Num_Eligible_Voters*NY),
    NY_count_win = count(New_Mem*NY)
  ) %>% 
  mutate(USA_count_NOwin = USA_count - USA_count_win,
         NY_count_NOwin = NY_count - NY_count_win,
         YEAR = year(year))

longer_table_count <- graph_table_count %>% 
  select(year, USA_work_win, NY_work_win, USA_work_NOwin, NY_work_NOwin) %>% 
  pivot_longer(c(USA_work_win, NY_work_win, USA_work_NOwin, NY_work_NOwin), names_to = "cat", values_to = "Workers") %>% 
  mutate(YEAR = year(year),
         geo = case_when(
           str_detect(cat, "USA") ~ "USA",
           str_detect(cat, "NY") ~ "NY"
         ),
         Win = case_when(
           str_detect(cat, "_win") ~ "Win",
           str_detect(cat, "_NOwin") ~ "No Win"
         ),
         geo = factor(geo, levels = c("USA", "NY")),
         Win = factor(Win)
         )


all_long_table_count <- graph_table_count %>% 
  select(year, USA_work_win, NY_work_win, USA_work, NY_work) %>% 
  pivot_longer(c(USA_work_win, NY_work_win, USA_work, NY_work), names_to = "cat", values_to = "Workers") %>% 
  mutate(YEAR = year(year),
         geo = case_when(
           str_detect(cat, "USA") ~ "USA",
           str_detect(cat, "NY") ~ "NY"
         ),
         Win = ifelse(str_detect(cat, "_win"), "Win", "All"),
         geo = factor(geo, levels = c("USA", "NY")),
         Win = factor(Win))


```

# Plots

## Setup

```{r eval=TRUE}

#remove.packages("Rttf2pt1")
#remotes::install_version("Rttf2pt1", version = "1.3.8")
#extrafont::font_import()

library(hrbrthemes)
library(scales)

library("ggthemes")
library(extrafont)

```

## Bars

```{r eval = FALSE}

worker_plot <- ggplot(data = longer_table, aes(x = YEAR, y = Workers, group = factor(geo), fill = factor(Win))) +
  geom_col(position = "dodge")

plot(worker_plot)

```

### One that Works

```{r eval = TRUE}

worker_plot2 <- ggplot() +
  geom_col(data = subset(all_long_table, Win == "All"), mapping = aes(x = YEAR, y = Workers, group = geo), position = "dodge", fill = "#4477AA", colour = "#fbf7ea") +
  geom_text(data = subset(all_long_table, Win == "All" & YEAR == 2019), mapping = aes(x = YEAR, y = Workers, group = geo, label = geo), position = position_dodge(width = 0.9), vjust = -1) +
  geom_col(data = subset(all_long_table, Win == "Win"), mapping = aes(x = YEAR, y = Workers, group = geo, fill = Win), position = "dodge", colour = "#fbf7ea") +
  scale_fill_manual(values = c("#CC6677")) +
#  geom_text(data = subset(all_long_table, Win == "Win" & YEAR == 2019), mapping = aes(x = YEAR, y = Workers, group = geo, label = Win), position = position_dodge(width = 1), vjust = -1) + 
  labs(x= element_blank(), 
       y= element_blank(),
       fill = element_blank(),
       title="NLRB Elections",
       subtitle="Number of Workers in NLRB Elections, by Year of Filing") + 
  #coord_cartesian(ylim = c(0,140000)) +
  scale_y_comma(limits = c(0,139000)) +
  theme_ipsum(grid = "Y",
              plot_margin = margin(10, 10, 10, 10),
              plot_title_margin = 5,
              subtitle_margin = 5) + 
  update_geom_font_defaults()

plot(worker_plot2)

```

```{r eval = FALSE}

worker_bars <- ggplot() +
  geom_col(data = graph_table, mapping = aes(x = YEAR, y = ))

```

### Save

```{r eval=TRUE}

ggsave(here("LMIS_src/graphs/nlrb_graph.png"), plot = worker_plot2, width = 7, height = 5, bg = "#fbf7ea")

# ggsave(here("LMIS_src/graphs/occupation_graph.png"), plot = occ_plot, bg = "#ebd999")

  # BG Colors: #ebd999, (lighter one) #fbf7ea

# ggsave(here("LMIS_src/graphs/occupation_graph.pdf"), plot = occ_plot, device = cairo_pdf())

```

## Doesn't work, with Loss/Pending

```{r eval = FALSE}

worker_plot3 <- ggplot() +
  geom_col(data = subset(all_long_table, Win == "All"), mapping = aes(x = YEAR, y = Workers, group = geo), position = "dodge", fill = "#4477AA", colour = "#fbf7ea") +
  geom_text(data = subset(all_long_table, Win == "All" & YEAR == 2019), mapping = aes(x = YEAR, y = Workers, group = geo, label = geo), position = position_dodge(width = 0.9), vjust = -1) +
  geom_col(data = subset(all_long_table, Win == "Win"), mapping = aes(x = YEAR, y = Workers, group = geo, fill = Win), position = "dodge", colour = "#fbf7ea") +
  scale_fill_manual(values = c("#CC6677")) +
#  geom_text(data = subset(all_long_table, Win == "Win" & YEAR == 2019), mapping = aes(x = YEAR, y = Workers, group = geo, label = Win), position = position_dodge(width = 1), vjust = -1) + 
  labs(x= element_blank(), 
       y= element_blank(),
       fill = element_blank(),
       title="NLRB Elections",
       subtitle="Number of Workers in NLRB Elections, by Year of Filing") + 
  #coord_cartesian(ylim = c(0,140000)) +
  scale_y_comma(limits = c(0,139000)) +
  theme_ipsum(grid = "Y",
              plot_margin = margin(10, 10, 10, 10),
              plot_title_margin = 5,
              subtitle_margin = 5) + 
  update_geom_font_defaults()

plot(worker_plot3)

```

### Save

```{r eval=FALSE}

ggsave(here("LMIS_src/graphs/nlrb_graph2.png"), plot = worker_plot3, width = 7, height = 5, bg = "#fbf7ea")

# ggsave(here("LMIS_src/graphs/occupation_graph.png"), plot = occ_plot, bg = "#ebd999")

  # BG Colors: #ebd999, (lighter one) #fbf7ea

# ggsave(here("LMIS_src/graphs/occupation_graph.pdf"), plot = occ_plot, device = cairo_pdf())

```

# Line

```{r eval=FALSE}

worker_line <- ggplot() +
  geom_line(data = graph_table, mapping = aes(x = YEAR, y = US_work), colour = "#4477AA", size = 2, linetype = 1) +
  geom_line(data = graph_table, mapping = aes(x = YEAR, y = US_work_win), colour = "#4477AA", size = 2, linetype = 2) +
  geom_line(data = graph_table, mapping = aes(x = YEAR, y = NY_work), colour = "#CC6677", size = 2, linetype = 1) +
  geom_line(data = graph_table, mapping = aes(x = YEAR, y = NY_work_win), colour = "#CC6677", size = 2, linetype = 2) +
  labs(x= element_blank(), y="% Unionization",
       title="Unionization by Occupation",
       subtitle="Comparing Unionization by Occupational Groups, 2019-2022") + 
  coord_cartesian(ylim = c(0,150000)) +
  theme_ipsum(grid = "Y")
  
  
plot(worker_line)

```

# Export Tables

```{r eval=FALSE}

write_csv(graph_table, "NLRB_table.csv")

```

# New Bar

## Loss/Pending

```{r eval = TRUE}

all_long_table2 <- graph_table %>% 
  select(year, USA_work_win, NY_work_win, USA_work, NY_work) %>% 
  pivot_longer(c(USA_work_win, NY_work_win, USA_work, NY_work), names_to = "cat", values_to = "Workers") %>% 
  mutate(YEAR = year(year),
         geo = case_when(
           str_detect(cat, "USA") ~ "USA",
           str_detect(cat, "NY") ~ "NY State"
         ),
         Win = ifelse(str_detect(cat, "_win"), "Win", "Loss/Pending"),
         geo = factor(geo, levels = c("USA", "NY State")),
         Win = factor(Win))

worker_plot_loss <- ggplot() +
  geom_col(data = subset(all_long_table2, Win == "Loss/Pending"), mapping = aes(x = YEAR, y = Workers, group = geo, fill = Win), position = "dodge", colour = "#fbf7ea") +
  geom_text(data = subset(all_long_table, Win == "All" & YEAR == 2019), mapping = aes(x = YEAR, y = Workers, group = geo, label = geo), position = position_dodge(width = 0.9), vjust = -1) +
  geom_col(data = subset(all_long_table2, Win == "Win"), mapping = aes(x = YEAR, y = Workers, group = geo, fill = Win), position = "dodge", colour = "#fbf7ea") +
  scale_fill_manual(values = c("#4477AA", "#CC6677")) +
#  geom_text(data = subset(all_long_table2, Win == "Win" & YEAR == 2019), mapping = aes(x = YEAR, y = Workers, group = geo, label = Win), position = position_dodge(width = 1), vjust = -1) + 
  labs(x= element_blank(), 
       y= element_blank(),
       fill = element_blank(),
       title="NLRB Elections",
       subtitle="Number of Workers in NLRB Elections, by Year of Filing") + 
  #coord_cartesian(ylim = c(0,140000)) +
  scale_y_comma(limits = c(0,139000)) +
  theme_ipsum(grid = "Y",
              plot_margin = margin(10, 10, 10, 10),
              plot_title_margin = 5,
              subtitle_margin = 5) + 
  update_geom_font_defaults()

plot(worker_plot_loss)

```

### Save

```{r eval=TRUE}

ggsave(here("LMIS_src/graphs/nlrb_graph_loss.png"), plot = worker_plot_loss, width = 7, height = 5, bg = "#fbf7ea")

# ggsave(here("LMIS_src/graphs/occupation_graph.png"), plot = occ_plot, bg = "#ebd999")

  # BG Colors: #ebd999, (lighter one) #fbf7ea

# ggsave(here("LMIS_src/graphs/occupation_graph.pdf"), plot = occ_plot, device = cairo_pdf())

```