---
title: "CPS_QP_Clean"
author: "Sam Neylon"
date: "July 21, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#### Using "here" to keep project directory as working directory
#### NOTE: Rmarkdown will try to make the file it is inside the working directory, but "here" should fix it.

library(here)
library(tidyverse)

```


# CPS Cleaning

I am using IPUMS CPS downloads (see codebook in data file where original download is kept). 

* Data set info:
* IPUMS CPS
* Time period: Bsic Monthly from 01-1990 to 12-2019

I only want observations which were included in the earner study - the outgoing groups. This should replicate the NBER MORG files, giving me a large n sample of the population.

I think that this code will just turn my raw files into a MORG data set. I will create a project specific markdown for my Qualifying Paper project.

# Today's Notebook

## (07-21-2021)

Goals for today:

* Import CPS csv's
* Extract only outgoing groups


# Import

## Import CSV

```{r echo=FALSE, warning=FALSE}

CPS_Master <- read_csv()


```




